---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 300,
  comment = "#>",
  fig.align = "center",
  fig.path = "man/figures/README-",
  fig.width = 15
)
```

# ðŸ—º usmap

<!-- badges: start -->
[![CRAN](http://www.r-pkg.org/badges/version/usmap?color=blue)](https://cran.r-project.org/package=usmap)
[![Downloads](http://cranlogs.r-pkg.org/badges/grand-total/usmap)](https://cran.r-project.org/package=usmap)
[![Build Status](https://img.shields.io/endpoint.svg?url=https%3A%2F%2Factions-badge.atrox.dev%2Fpdil%2Fusmap%2Fbadge%3Fref%3Dmaster&style=popout&label=build)](https://actions-badge.atrox.dev/pdil/usmap/goto?ref=master)
[![codecov](https://codecov.io/gh/pdil/usmap/branch/master/graph/badge.svg)](https://app.codecov.io/gh/pdil/usmap)
<!-- badges: end -->

```{r header, echo=FALSE, fig.height=8}
library(usmap)
library(ggplot2)

# City population county map (West Region) ####
citypop_t <- usmap_transform(citypop)
citypop_t_west <- citypop_t[(citypop_t$abbr %in% .west_region) & (citypop_t$abbr != "AK") & (citypop_t$abbr != "HI"), ]

west_county_citypop_map <-
  plot_usmap("counties", include = .west_region, exclude = c("AK", "HI"), color = "blue", fill = "lightblue") +
  geom_sf(data = citypop_t_west, aes(size = city_pop), color = "red", alpha = 0.7) +
  scale_size_continuous(range = c(2, 15), guide = "none") +
  ggrepel::geom_label_repel(
    data = citypop_t_west,
    aes(label = most_populous_city, geometry = geometry),
    size = 5, alpha = 0.8,
    label.r = unit(0.5, "lines"), label.size = 0.5,
    segment.color = "red", segment.size = 1,
    stat = "sf_coordinates", seed = 1002,
    max.overlaps = 20
  )

# Population by state with labels ####
state_pop_map_labeled <-
  plot_usmap(data = statepop, values = "pop_2015", labels = TRUE) +
  scale_fill_continuous(low = "white", high = "red", guide = "none")

# Blank county map (Alaska) ####
ak_county_map <-
  plot_usmap("counties", include = "AK", color = "red", fill = "#fffdcf")

# Poverty by county (South) ####
south_pov_map <-
  plot_usmap("counties", data = countypov, values = "pct_pov_2014",
             include = .south_region, color = "white", size = 0) +
  scale_fill_continuous(low = "darkgreen", high = "yellow", guide = "none")

# Poverty percentage by county ####
county_pov_map <-
  plot_usmap(data = countypov, values = "pct_pov_2014", size = 0.2) +
  scale_fill_continuous(low = "blue", high = "yellow", guide = "none")

# Rivers map ####
rivers_t <- usmap_transform(usrivers)

rivers_map <- plot_usmap("counties", color = "gray80") +
  geom_sf(data = rivers_t, aes(linewidth = Shape_Length), color = "blue") +
  scale_linewidth_continuous(range = c(0.3, 1.5), guide = "none")

# Combine plots ####
cowplot::plot_grid(
  west_county_citypop_map,
  state_pop_map_labeled,
  ak_county_map,
  south_pov_map,
  county_pov_map,
  rivers_map,
  nrow = 2
)
```

## Purpose

Typically in R it is difficult to create nice US [choropleths](https://en.wikipedia.org/wiki/Choropleth_map) that include Alaska and Hawaii. The functions presented here attempt to elegantly solve this problem by manually moving these states to a new location and providing a simple features ([`sf`](https://github.com/r-spatial/sf)) object for mapping and visualization. This allows the user to easily add spatial data or features to the US map.

## Shape Files

The shape files that we use to plot the maps in R are located in the [`usmapdata`](https://github.com/pdil/usmapdata) package. These are generated from the [US Census Bureau cartographic boundary files](https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html). Maps at both the state and county levels are included for convenience.

#### Update History

| Date              | `usmap` version | Shape File Year |                                                  Link                                                  |
|------------------|:----------------:|:----------------:|:----------------:|
| (unreleased)      |      0.7.0      |      2022       | [ðŸ”—](https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.2022.html) |
| February 27, 2022 |      0.6.0      |      2020       | [ðŸ”—](https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.2020.html) |
| June 3, 2018      |      0.3.0      |      2017       |  [ðŸ”—](https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2017.html)  |
| January 29, 2017  |      0.1.0      |      2015       |  [ðŸ”—](https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2015.html)  |

## Installation

ðŸ“¦ To install from CRAN (recommended), run the following code in an R console:

```{r install, eval=FALSE}
install.packages("usmap")
```

### Developer Build

âš ï¸ The developer build may be unstable and not function correctly, use with caution.

To install the package from this repository, run the following code in an R console:

```{r install_dev, eval=FALSE}
install.package("devtools")
devtools::install_github("pdil/usmap")
```

This method will provide the most recent developer build of `usmap`.

To begin using `usmap`, import the package using the `library` command:

```{r load}
library(usmap)
```

## Documentation

To read the package vignettes, which explain helpful uses of the package, use `vignette`:

```{r vignettes, eval=FALSE}
vignette(package = "usmap")
vignette("usmap1", package = "usmap") # 1. Introduction
vignette("usmap2", package = "usmap") # 2. Mapping the US
vignette("usmap3", package = "usmap") # 3. Advanced Mapping
```

For further help with this package, open an [issue](https://github.com/pdil/usmap/issues) or ask a question on Stack Overflow with the [usmap tag](https://stackoverflow.com/questions/tagged/usmap).

## Features

### Map Plots

- Plot US maps

```{r plots, fig.align='center'}
states <- plot_usmap("states")
counties <- plot_usmap("counties")

cowplot::plot_grid(states, counties, nrow = 1)
```


- Display only certain states, counties, or regions

```{r more_plots}
library(ggplot2)

mt <- plot_usmap("states", include = .mountain, labels = TRUE)

fl <- plot_usmap("counties", data = countypov, values = "pct_pov_2014", include = "FL") +
  scale_fill_continuous(low = "green", high = "red", guide = "none")

ne <- plot_usmap("counties", data = countypop, values = "pop_2015", include = .new_england) +
  scale_fill_continuous(low = "blue", high = "yellow", guide = "none")

cowplot::plot_grid(mt, fl, ne, nrow = 1)
```


- Transform and add spatial data to map

```{r sf_plot}
library(ggplot2)

# Transform included `usrivers` data set
rivers_transformed <- usmap_transform(usrivers)

river_map <- plot_usmap("counties", color = "gray80") +
  geom_sf(data = rivers_transformed, aes(linewidth = Shape_Length), color = "blue") +
  scale_linewidth_continuous(range = c(0.3, 1.5), guide = "none")

# Transform included `earthquakes` data set
eq_transformed <- usmap_transform(earthquakes)

earthquake_map <- plot_usmap() +
  geom_sf(data = eq_transformed, aes(size = mag), color = "red", alpha = 0.25) +
  scale_size_continuous(guide = "none")

cowplot::plot_grid(river_map, earthquake_map, nrow = 1)
```


### Map Data

- Obtain map data with certain region breakdown
```{r state_map}
us_map(regions = "states")
```

```{r county_map}
us_map(regions = "counties")
```


### FIPS Codes

- Look up FIPS codes for states and counties

```{r fips}
fips("New Jersey")

fips(c("AZ", "CA", "New Hampshire"))

fips("NJ", county = "Mercer")

fips("NJ", county = c("Bergen", "Hudson", "Mercer"))
```


- Retrieve states or counties with FIPS codes

```{r fips_info}
fips_info(c("34", "35"))

fips_info(c("34021", "35021"))
```


- Add FIPS codes to data frame

```{r fips_merge, message=FALSE}
library(dplyr)

data <- data.frame(
  state = c("NJ", "NJ", "NJ", "PA"),
  county = c("Bergen", "Hudson", "Mercer", "Allegheny")
)

data %>% rowwise %>% mutate(fips = fips(state, county))
```


## Additional Information

### Coordinate System

`usmap` uses the [US National Atlas Equal Area](https://epsg.io/9311) coordinate system:

<details>
  <summary><code>sf::st_crs(9311)</code></summary>

  ```{r crs, echo=FALSE}
  sf::st_crs(9311)
  ```
</details>

This [coordinate reference system (CRS)](https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf) can also be obtained with `usmap::usmap_crs()`.

## Acknowledgments

The code used to generate the map files was based on this blog post by [Bob Rudis](https://github.com/hrbrmstr): [Moving The Earth (well, Alaska & Hawaii) With R](https://rud.is/b/2014/11/16/moving-the-earth-well-alaska-hawaii-with-r/)
